---
editor_options:
  chunk_output_type: console
output:
  html_document: default
  pdf_document: default
  word_document: default
---
##GitHub website: https://github.com/klbrundage 
##01-Assignment_Brundage_K

```{r}
##Printing/Showing the Statement "Hello, World"
print("Hello, World")
```


```{r load_data}
## Load in the data file sc - College Data File
load("college.Rdata")
```


```{r look_at_data}
## What does this data look like? Look at the first few rows, first few variables
sc

## This is commented as you might not always want to run it. Delete the "#" sign below in order to run the command. 
View(sc)
```


```{r Mean Cost College}
##Obtains the Mean Cost of College
sc%>%summarize(mean_cost=mean(costt4_a,na.rm=TRUE))
```


##Stretch Items

##1: Calculate the average earnings for individuals at the most selective colleges, then compare that with individuals at the least selective colleges in the dataset.

```{r combining commands}
## Gives the average earnings for individuals at the most selective colleges, then compare that with individuals at the least selective colleges
sc%>%filter(adm_rate<.1)%>%summarize(mean_earnings1=mean(md_earn_wne_p6,na.rm = TRUE))

##Gives the average earnings for non-selective schools from the College Data File
sc%>%filter(adm_rate>.3)%>%summarize(mean_earnings2=mean(md_earn_wne_p6,na.rm = TRUE))
```

##2: Find a way to determine whether colleges with very high SAT scores tend to be larger or smaller than colleges with low SAT scores.

```{r SAT Scores}
## Left Side: Colleges with high SAT Scores and the average number of undergrads at those colleges
sc%>%filter(sat_avg>1400)%>%summarize(mean_high=mean(ugds,na.rm = TRUE))

##Right Side: Colleges with low SAT Scores and the average number of undergrads at those Colleges
sc%>%filter(sat_avg<900)%>%summarize(mean_low=mean(ugds,na.rm = TRUE))
```

##3:Plot the relationship between cost and debt. What do you see? Does this surprise you?

##Answer to Q3:  There is a positive relationship between cost and debt.  The higher average cost for attending an instiution the higher the average debt one will incur. No, this does not surprise me, especially since I work in higher education and have experience in Financial Aid as well.  I have seen this trend over the past twenty-five years.

```{r Plot Cost v Debt}
## Plots the relationship between cost of College and the debt from the College Data File
cd<-ggplot(data=sc,mapping=aes(x=debt_mdn, y=costt4_a))+
  geom_point()
  geom_smooth()
  cd<-cd+xlab("Mean Debt")+ylab("Cost")
  cd
```

##4:Now, provide separate plots for cost and debt by control of the institution.

```{r Separate Plots}
## Left Side: plot showing the debt based on the control indicator of each instiution in the College Data File
cd1<-ggplot(data=sc)+
  geom_smooth(mapping=aes(x=debt_mdn, y=control))
cd1<-cd1+xlab("Mean Debt")+ylab("Control")
cd1

## Right Side: plot showing the cost based on the control indicator of each institution in the College Data File
cd2<-ggplot(data=sc)+
  geom_smooth(mapping=aes(x=debt_mdn, y=control))
cd2<-cd2+xlab("Mean Debt")+ylab("Control")
cd2
```
