---
title: "01-Assignment_Brundage"
GitHub: https://github.com/klbrundage
output:
  pdf_document: default
  html_document: default
  word_document: default
---

```{r}
##Printing/Showing the Statement "Hello, World"
print("Hello, World")
```

```{r load_data}
## Load in the data file sc - College Data File
load("college.Rdata")
```


```{r summarize}
## Gives the median cost of college based on the College Data File
sc%>%summarize(mean_cost=mean(costt4_a,na.rm=TRUE))
```

##Stretch Items

##1: Calculate the average earnings for individuals at the most selective colleges, then compare that with individuals at the least selective colleges in the dataset.

```{r combining commands}
## Gives the average earnings for very selective schools from the College Data File
sc%>%filter(adm_rate<.1)%>%summarize(mean_earnings1=mean(md_earn_wne_p6,na.rm=TRUE))

## Gives the average earnings for non-selective shcools from the College Data File
sc%>%filter(adm_rate>.3)%>%summarize(mean_earnings2=mean(md_earn_wne_p6,na.rm=TRUE))
```


##2: 2.	Find a way to determine whether colleges with very high SAT scores tend to be larger or smaller than colleges with low SAT scores.

```{r}
## Left Side: Colleges with high SAT Scores and the average number of undergrads at those colleges
sc%>%filter(sat_avg>1400)%>%summarize(mean_high=mean(ugds,na.rm=T))

##Right Side: Colleges with low SAT Scores and the average number of undergrads at those colleges
sc%>%filter(sat_avg<900)%>%summarize(mean_low=mean(ugds,na.rm=T))
```

##3:Plot the relationship between cost and debt. What do you see? Does this surprise you?

##Answer to Q3:  There is a positive relationship between cost and debt.  The higher average cost for attending an instiution the higher the average debt one will incur. No, this does not surprise me, especially since I work in higher education and have experience in Financial Aid as well.  I have seen this trend over the past twenty-five years.

```{r}
##Plots the relationship between cost of College and the debt from the College data file
cd<-ggplot(data=sc, mapping=aes(x=debt_mdn, y=costt4_a))+
    geom_point()+
    geom_smooth()
cd<-cd+xlab("Mean Debt")+ylab("Cost")
cd
```

##4:Now, provide separate plots for cost and debt by control of the institution.

```{r}
#Left Side: plot showing the debt based on the control indicator of each institution in the College Data File
cd1<-ggplot(data=sc)+
    geom_smooth(mapping=aes(x=debt_mdn, y=control))
cd1<-cd1+xlab("Mean Debt")+ylab("Control")
cd1

#Right Side: plot showing the cost based on the control indicator of each instituion in the College Data File
cd2<-ggplot(data=sc)+
    geom_smooth(mapping=aes(x=costt4_a, y=control))
cd2<-cd2+xlab("Cost")+ylab("Control")
cd2

```

