---
title: "02-Assignment_Brundage"
output:
  pdf_document: default
  word_document: default
  html_document: default
GitHub: https://github.com/klbrundage
Date: May 23, 2019
Author: Kelley Brundage
---

```{r global_options, include = FALSE}
knitr::opts_chunk$set(message=FALSE, 
tidy.opts=list(width.cutoff=60)) 
```

#Libraries and Data Files
```{r, warning=F}
##Load libraries in order to successfully run the code below
library(tidyverse)
library(plotly)
library(Metrics)
```


```{r data}
##Load the County Level Dataset
load("pd.Rdata")
```

```{r explain_date}
## Full explanation of data in codebook
load("pd_lab_explain.Rdata")

#or use View
#View(lab_explain)
```

##Assignment 02

#For this assignment, you'll be working with the county level dataset to predict a home ownership rates using conditional means. You'll need to select the county-level characteristics that you think might be related to home ownership rates. Please complete the following steps:

--------------------------------------------------------
#Question 1: Calculate the mean of the outcome.
```{r mean}
##This code creates the mean of the home ownership value within the pd.data file and gives the output of the mean
pd%>%summarize(mean_homeown=mean(homeown_rate,na.rm=TRUE))
```

--------------------------------------------------------
#Question 2: Use your mean as a prediction: Create a new variable that consists of the mean of the outcome.

```{r new variable for prediction}
##Creation of a new Variable, homeownership, that shows the mean of median household income which has been labled homeown_rank.  No output will be relected as this is being added to the pd.data file as a new field.

pd<-pd%>%mutate(homeown_rank=rank(median_hh_inc))
```

```{r place new variable in dataset}
##Unconditional Average as a Predictor
##This is also creating a new variable that will be added to the pd.data file as a new field reflecting the mean of median household income but this variable is named mean_home.

pd <- pd%>%mutate(mean_home=mean(median_hh_inc,na.rm = T))
```

```{r Plot Average with Actual}
##Plotting
##Here we are plotting the median household income compared to the rank of median household income.
gg<-ggplot(data=pd,aes(y=median_hh_inc,x=homeown_rank,color="Actual"))
gg<-gg+geom_point(alpha=.5,size=.5)
gg<-gg+geom_point(aes(y=mean_home,x=homeown_rank,
                  color="Predicted: Unconditional Mean"),
                  size=.5)
gg<-gg+xlab("Rank of Median Household Income")+ylab("Median Household Income")
gg<-gg+scale_color_manual(name="Type",
                          values=c("Actual"="black",
                          "Predicted: Unconditional Mean"="blue")
                          )
gg<-gg+theme(legend.position="bottom")

gg

##Save for later

gg2<-gg
```

--------------------------------------------------------
#Question 3: Calculate a summary measure of the errors for each observation-the difference between your prediction and the outcome.

```{r error}
##The code below will calculate a new variable
pd<-pd%>%mutate(e2=median_hh_inc-mean_home)
```

```{r RMSE}
##The Root Mean Squared Error (RMSE) shows the difference in median household income versus the mean of home predictor created in question #2 above
rmse_uncond_mean<-rmse(pd$median_hh_inc, pd$mean_home)

rmse_uncond_mean
```

--------------------------------------------------------
#Question 4: Calculate the mean of the outcome at levels of a predictor variable.
```{r Conditional Mean Single}
##Conditional Average across a single variable which is called the home_value
pd<-pd%>%mutate(home_value=ntile(median_home_val,4))

table(pd$home_value)##this code will create a table of the new home_value variable

pd<-pd%>%group_by(home_value)%>% ## This code will take the home_value variable and group it together by predictor
  mutate(pred_value=mean(median_home_val))%>%  ##The mutate code will calculate mean at each level of predictor
  ungroup()%>%  ##This code will ungroup the data
  mutate(pred_value_rank=rank(pred_value,ties.method="random")) #This code will result in the rank by prediction, with ties sorted randomly

```

--------------------------------------------------------
#Question 5: Use these conditional means as a prediction: for every county, use the conditional mean to provide a ''best guess" as to that county's level of the outcome.
```{r quartiles}
##Create a variable for quartiles for home ownership by creating a varable named homeown_rate_level.  The results will not show in this code as it is added the field to the pd.data file and putting the data together in 4 bins
pd <- pd%>%mutate(homeown_rate_level=ntile(homeown_rate,4))
```

```{r mean of predictor}
##Calculate the mean at each level of the predictor, which is the pred_inc_homeown variable based on the homeown_rate_level
pd%>%group_by(homeown_rate_level)%>% summarise(pred_inc_homeown=mean(median_hh_inc))
```

The error term for our prediction using unconditional means will be stored in the variable $e1$. This variable will be equal to the actual value of per capita income `percapinc.2010` minues the mean value of per capita income `mean_percapinc.2010`. 
```{r mean of each predictor}
pd<-pd%>%group_by(home_value,homeown_rate_level)%>% ## This code will take the home_value variable and group it together by predictor
  mutate(pred_value_and_homeown=mean(median_hh_inc))%>% ##The mutate code will calculate mean at each level of predictor
  ungroup()%>% ##This code will ungroup the data
  mutate(pred_value_and_homeown_rank=rank(pred_value_and_homeown,                                              ties.method="random"))##This code will result in the rank by prediction, with ties sorted randomly
```

--------------------------------------------------------
#Question 6:Calculate a summary measure of the error in your predictions.

```{r error_terms}
##Calculates the Root Mean Squared Error (RMSE)
rmse_cond_mean_two <- rmse(pd$median_hh_inc,pd$pred_value_and_homeown)
rmse_cond_mean_two
```

```{r round RMSE}
##This code will round the RMSE with no decimal places
round(rmse_cond_mean_two,0)
```

