---
title: "03-Assignment_BrundageK"
Date: "May 23, 2019"
GitHub: "https://github.com/klbrundage/LLO8200-Assignments"
output:
  pdf_document: default
  html_document: default
  word_document: default
Author: "Kelley Brundage"

---

```{r global_options, include = FALSE}
##This code does not show in the final document but will assist with definining the margin cutoff point and wraps the text to the next line.
knitr::opts_chunk$set(message=FALSE, 
tidy.opts=list(width.cutoff=60)) 
```

# *Setup for plotting conditional means*

We always start with a standard set of setup commands by loading the correct libraries. We will continue to work with `tidyverse`, and are adding `forcats` which helps us to deal with factor variables as well as `RColorBrewer' to handle colors.

```{r, warning=F}
##Load libraries in order to successfully run the code below
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(forcats)
library(RColorBrewer)
library(knitr)
```

## *Loading Data* 
```{r load data}
##The code below will oad the data file attrition and open the file so it can be viewed
load("attrition.Rdata")
View (at)
```

## **Question 1: Create a graph that shows the distribution of monthly income.**

```{r monthly income histogram graph}
##The code below creates a Histogram showing the distributin of monthly income from the attrition data file
gg<-ggplot(at,aes(x=MonthlyIncome))
gg<-gg+geom_histogram(binwidth = 20,fill="darkgreen")

## Labels for the Histogram Chart
gg<-gg+ylab("Count")+xlab("Monthly Income") ## x & y axis labels
gg<-gg+theme(axis.text.x = element_text(angle = 60, hjust = 1))
gg<-gg+ggtitle("Distribution of Monthly Income") ## Chart Title
gg
``` 

## **Question 2: Create a graph that shows the average level of monthly income by field of education.**

```{r Level of Income by Education}
##The code below creates a new variable at_sum for the Monthly Income field
at_sum<-at%>%group_by(MonthlyIncome, EducationField)%>%
  summarize(mincome=mean(MonthlyIncome))%>%arrange(-mincome)
  ## gives the mean of Monthly Income
at_sum
```

```{r Bar Chart for Mincome by Field of Ed}
##The code below creates a bar chart showing the average level of monthly income by field of education
gg<-ggplot(at_sum,aes(x=EducationField,y=mincome,
        fill=EducationField)) ##fills in the bars for each field of education
## Bar plot, with unstacked (dodge)
 gg<-gg+geom_bar(stat="identity",position="dodge")

##Code that will label the Bar Chart
gg<-gg+ylab("Mean of Monthly Income")+xlab("Education Field") ## x & y axis labels
gg<-gg+theme(axis.text.x = element_text(angle = 60, hjust = 1))
gg<-gg+ggtitle("Level of Monthly Income by Field of Education") ## Chart Title
gg
```

## **Question 3: Create another graph that shows the average level of monthly income by field of education and gender.**

```{r Level of Income by Education & Gender}
## Summarize by Monthly Income, Education Field and Gender
at_sum<-at%>%group_by(MonthlyIncome, EducationField, Gender)%>%
  summarize(mincome=mean(MonthlyIncome))%>%arrange(-mincome)
  ## gives the mean of Monthly Income
at_sum
```

```{r bar chart by Gender}
## The code below uses Faceting which means we are making multiple graphs with the same structure. 

## The code arranges positions based on Education Field, color based on gender, and then split the graphic by Monthly Income. 
gg<-ggplot(at_sum,aes(x=reorder(EducationField,mincome),
                      y=mincome,
                      fill=Gender))
## Bar plot, with unstacked (dodge)
 gg<-gg+geom_bar(stat="identity",position="dodge")
  ## Separate out by Gender
gg<-gg+facet_wrap(~Gender)
  ## Change orientation to sideways
gg<-gg+coord_flip()

##Code that adds lables to the bar chart
gg<-gg+ylab("Mean of Monthly Income")+xlab("Education Field")##lables the x & y axis
gg<-gg+theme(axis.text.x = element_text(angle = 60, hjust = 1))##assigns the theme to the chart
gg<-gg+ggtitle("Level of Monthly Income by Field of Education & Gender")##Title of the Chart
gg
```

## **Question 4: Create a graph that shows average levels of monthly income by field of education, gender and job level (scale of 1-5, highest ranked employees are 5)**

```{r Mincome by FoE, Gender & Job Level}
## Summarize by Monthly Income, Education Field, Gender & Job Level
at_sum<-at%>%group_by(MonthlyIncome, EducationField, Gender, JobLevel)%>%
  summarize(mincome=mean(MonthlyIncome))%>% ## gives the mean of Monthly Income
  ungroup()%>%
  arrange(mincome)
at_sum
```

```{r, echo=FALSE, results='hide', warning=FALSE, message=FALSE}
## The code below also uses Faceting which means we are making multiple graphs with the same structure. 

## The code arranges positions based on Education Field, color based on gender, and then split the graphic by Monthly Income. 

gg<-ggplot(at_sum,aes(x=fct_reorder(Gender,mincome),y=mincome))
gg<-gg+geom_bar(stat="identity",aes(fill=EducationField),position="dodge")
gg<-gg+facet_wrap(~Gender+JobLevel,ncol=4)

##Code that adds lables to the Bar Chart
gg<-gg+ylab("Mean of Monthly Income")+xlab("Gender")##labels the x & y axis
gg<-gg+theme(axis.text.x = element_text(angle = 60, hjust = 1))##assigns the theme to the chart
gg<-gg+ggtitle("Level of Monthly Income by Field of Education, Gender & Job Level")##Creates the Title for the Chart
gg
```
