---
title: "06-Assignment_BrundageK"
Date: May 28, 2019
GitHub: https://github.com/klbrundage/LLO8200-Assignments
output:
  pdf_document: default
  html_document: default
Author: Kelley Brundage
---

```{r setup, include=FALSE}
##This code allows the Knit function to still work even with errors 
knitr::opts_chunk$set(echo=TRUE,error=TRUE)
```

```{r global_options, include = FALSE}
##This code does not show in the final document but will assist with definining the margin cutoff point and wraps the text to the next line.
knitr::opts_chunk$set(message=FALSE, 
  tidy.opts=list(width.cutoff=60)) 

  my_pdf = function(file,width,height)
  {pdf(file, width=width, height=height,pointsize=12)}
```

# *Setup for Scatterplots - Loading the Libraries*

We always start with a standard set of setup commands by loading the correct libraries. We will continue to work with `tidyverse` and others and will add 'haven' and 'readxl' in order to ensure we have tidy data.

```{r, warning=F}
##Load libraries in order to successfully run the code below
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(dplyr)
library(evaluate)
library(forcats)
library(formatR)
library(haven)
library(knitr)
library(ModelMetrics)
library(modelr)
library(readxl)
library(tibble)
```

Complete the following steps using the .Rdata or file. I want you to use the ggplot library and commands for all plots. 

Use the els data to complete the following steps.

The ELS (Educational Longitudinal Survey) dataset is called `els`.

```{r Load ElS Datasets}
load("els.RData")
```

#**Question#1: Choose a dependent variable and plot that dependent variable as a function of a continuous independent variable.**

```{r Plotting Spending v Income}
#Plot SES vs. Reading Scores
g1<-ggplot(data=els,aes(x=byses1,y=bynels2r))
g1<-g1+geom_point() # Add points at x and y

g1
```

#**Question#2: If you need to, adjust the scale of either the dependent variable or independent variable to make more sense.**

```{r Adjusting Scale in Plot}
#Plot SES vs. Reading Scores
g1<-ggplot(data=els,aes(x=byses1,y=bynels2r))
g1 <- g1+geom_point(alpha=.5,size=.25) 
g1 <- g1+scale_y_continuous(breaks = seq(10,50, by=5))
g1 <- g1+scale_color_discrete()
g1
```

#**Question#3: Add a line of best fit to your graphic.**

```{r Q3: LOESS Line}
#This will assist with getting the massive dataset down to something a bit more reasonable to work with
els_sum<-els%>%
  mutate(ses_rank=percent_rank(byses1)*100)%>%
  mutate(ses_rank_r=round(ses_rank))%>% 
      #this command rounds so there are no decimal points
  group_by(ses_rank_r)%>% #group by the new rounded variable
  summarize(test_mean=mean(bynels2r,na.omit=TRUE)) #summarize the dataset

g1a<-ggplot(els_sum,aes(x=ses_rank_r,y=test_mean))
g1a<-g1a+geom_point(alpha=.5,size=.25)

g1a<-g1a+geom_smooth() # will default to the LOESS method

g1a

```

```{r Q3: Linear Regression Line}
#This will assist with getting the massive dataset down to something a bit more reasonable to work with
els_sum<-els%>%
  mutate(ses_rank=percent_rank(byses1)*100)%>%
  mutate(ses_rank_r=round(ses_rank))%>% 
      #this command rounds so there are no decimal points
  group_by(ses_rank_r)%>% #group by the new rounded variable
  summarize(test_mean=mean(bynels2r,na.omit=TRUE)) #summarize the dataset

g1b<-ggplot(els_sum,aes(x=ses_rank_r,y=test_mean))
g1b<-g1b+geom_smooth(method = "lm") # will default to the LOESS method

g1b

```


#**Question#4: Clean up the labeling and description of the graphic so that it could be presented to a broader audience.**

```{r Q4: LOESS Line}
#The code below assist with getting the ELS dataset down to something a bit more reasonable to work with for plotting and comparison
els_sum<-els%>%
  mutate(ses_rank=percent_rank(byses1)*100)%>%
  mutate(ses_rank_r=round(ses_rank))%>% 
      #this command rounds so there are no decimal points
  group_by(ses_rank_r)%>% #group by the new rounded variable
  summarize(test_mean=mean(bynels2r,na.omit=TRUE)) #summarize the dataset

#The code below creates the LOESS chart
g1a<-ggplot(els_sum,aes(x=ses_rank_r,y=test_mean))
g1a<-g1a+geom_point(shape=2,alpha=.5,size=.5)#shape, style and size of points
g1a<-g1a+geom_smooth() # will default to the LOESS method

g1a<-g1a+ylab("Reading Test Scores")+xlab("SES Rank")#Labels for x & y axis
g1a<-g1a+theme_light() #theme with light grey lines and axes which direct more attention towards the data
g1a<-g1a+ggtitle("LOESS: Reading Scores Compared to SES")#Chart Title

g1a

```

```{r Q4: Linear Regression Line}
#The code below will assist with getting the ELS dataset down to something a bit more reasonable to work with for plotting and comparison
els_sum<-els%>%
  mutate(ses_rank=percent_rank(byses1)*100)%>%
  mutate(ses_rank_r=round(ses_rank))%>% 
      #this command rounds so there are no decimal points
  group_by(ses_rank_r)%>% #group by the new rounded variable
  summarize(test_mean=mean(bynels2r,na.omit=TRUE)) #summarize the dataset

#The code below creates the Linear Regression Chart
g1b<-ggplot(els_sum,aes(x=ses_rank_r,y=test_mean))
g1b<-g1b+geom_point(shape=2,alpha=.5,size=.5)#shape, style and size of points
g1b<-g1b+geom_smooth(method = "lm")#will plot a linear regression line

g1b<-g1b+ylab("Reading Test Scores")+xlab("SES Rank")#labels for x & y axis
g1b<-g1b+theme_light() #theme with light grey lines and axes which direct more attention towards the data
g1b<-g1b+ggtitle("Linear Regression: Reading Scores Compared to SES")#chart title

g1b

```

#**Question#5: In your .Rmd file, include a few sentences describing what the graphic above shows.**

The two graphs above in question #4 show that Socio-economic status increases as reading scores are predicted to increase.  When comparing the LOESS line to the Linear Regression line you see little difference in the position of the line or the plot points.  Due to this minute difference I would recommend using the Linear Regression Line since there does not seem to be a complicated curve in the data.  As a result the linear model shows the best possible result (line) through the data points.
