---
title: "07-Assignment_BrundageK"
Date: "June 16, 2019"
GitHub: "https://github.com/klbrundage/LLO8200-Assignments"
output:
  pdf_document: 
    fig_caption: yes
    fig_height: 3
    fig_width: 5
    keep_tex: yes
    number_sections: yes
    toc: yes
  word_document: 
    fig_caption: yes
    fig_height: 3
    fig_width: 5
    keep_md: yes
    toc: yes
  html_document: 
    fig_caption: yes
    fig_height: 3
    fig_width: 5
    keep_md: yes
    number_sections: yes
    toc: yes
Author: "Kelley Brundage"
---

```{r setup, include=F, error=F, fig.show='asis'}
##This code allows the Knit function to still work even with errors 
knitr::opts_chunk$set(echo=TRUE,error=TRUE)
```

```{r global_options, include = FALSE, message=F, warning=F,error=F, fig.show='asis'}
##This code does not show in the final document but will assist with definining the margin cutoff point and wraps the text to the next line.
knitr::opts_chunk$set(fig.path = "Figs/", message=FALSE, warning = F,
  tidy.opts=list(width.cutoff=60)) 

  my_pdf = function(file,width,height)
  {pdf(file, width=width, height=height,pointsize=12)}
```

# *Setup for Scatterplots - Loading the Libraries*

We always start with a standard set of setup commands by loading the correct libraries. We will continue to work with `tidyverse` and others and will add 'acs', 'lubridate', 'noncensus', 'rvest' and 'tigris'.

```{r, eval=F, tidy=T, warning=F}
##Load libraries in order to successfully run the code below
knitr::opts_chunk$set(echo = TRUE)
suppressMessages(library(acs))
suppressMessages(library(dplyr))
suppressMessages(library(evaluate))
suppressMessages(library(forcats))
suppressMessages(library(formatR))
suppressMessages(library(ggplot2))
suppressMessages(library(haven))
suppressMessages(library(knitr))
suppressMessages(library(lubridate))
suppressMessages(library(ModelMetrics))
suppressMessages(library(modelr))
suppressMessages(library(noncensus))
suppressMessages(library(readxl))
suppressMessages(library(rvest))
suppressMessages(library(tibble))
suppressMessages(library(tidyverse))
suppressMessages(library(tigris))
```


For this assignment, I want you to use the Census ACS API to download and
plot data. 

#Question #1: Download data for all of the zip codes in Fulton county, GA on education levels.

```{r fips codes from Tigris package}
## Look up fips code for county
lookup_code("GA","Fulton") 
#[1] "The code for Georgia is '13' and the code for Fulton County is '121'."

state_fips<-"13"
county_stub<-"121"
```

Next, we'll combine the state and county fips into a single object

```{r Combine State & county Fips}
county_fips<-paste0(state_fips,county_stub)
```

```{r Census Dataset}

# Get the census.gov dataset that matches all zip codes to cities, counties and states. 
county_to_zip<-read_csv("http://www2.census.gov/geo/docs/maps-data/data/rel/zcta_county_rel_10.txt")
save(county_to_zip,file="county_to_zip.Rdata") #saves this data to a new file named county_to_zip.Rdata

#Renames the County_to_zip to make it easier to work with
names(county_to_zip)<-tolower(names(county_to_zip))

#the code below returns Just zip codes in selected county - Fulton County, GA
county_to_zip<-county_to_zip%>%
  filter(state==as.numeric(state_fips),county==county_stub)%>%
  select(zcta5,state,county)

#creates a value that gives the list of zip codes in Fulton County, GA
ziplist<-county_to_zip$zcta5

# pulls in the City names affilited with each zip code
data(zip_codes)

city_zip<-zip_codes%>%filter(zip%in%ziplist)%>%select(zip,city)

#Arrange the cip codes in order by number
city_zip<-city_zip%>%arrange(as.numeric(zip))
```

```{r ACS Key & Fulton County Ed Level Data}

acs_key<-"2a43c63586cccde1824cd035269dfb621b16146c"

#List of tables: https://www.census.gov/programs-surveys/acs/technical-documentation/summary-file-documentation.html under, 1-year appendices
# b15001: education of pop over 18, by sex 
# b19001: household income over last 12 months

api.key.install(acs_key, file = "key.rda")

select_zip<-geo.make(zip.code=ziplist)

county_educ=acs.fetch(geography=select_zip,
                      endyear=2016,
                      table.number="B15001",
                      col.names="pretty",verbose=T)
save(county_educ,file="county_educ_ga.Rdata") #This may take a long time to run as it is a lot of data to parse and run
acs.colnames(county_educ)
```


#Question #2: Compute the proportion of the population that has a bachelor’s degree or above by zip code.
```{r Proportion of Pop with Bach Degree or Higher}

#The code below will compute the proportion of the LA county population that has a bachelor's degree or higher by zip code
prop_bdegplus<-divide.acs(numerator=(county_educ[,9]+ #Male: Bach Deg age 18-24
                            county_educ[,10]+ #Male: Grad/Prof Deg age 18-24
                            county_educ[,17]+ #Male: Bach Deg age 25-34
                            county_educ[,18]+ #Male: Grad/Prof Deg age 25-34
                            county_educ[,25]+ #Male: Bach Deg age 35-44
                            county_educ[,26]+ #Male: Grad/Prof Deg age 35-44
                            county_educ[,33]+ #Male: Bach Deg age 45-64
                            county_educ[,34]+ #Male: Grad/Prof Deg age 45-64
                            county_educ[,41]+ #Male: Bach Deg age 65+
                            county_educ[,42]+ #Male: Grad/Prof Deg age 65+
                            county_educ[,50]+ #Female: Bach Deg age 18-24
                            county_educ[,51]+ #Female: Grad/Prof Deg age 18-24
                            county_educ[,58]+ #Female: Bach Deg age 25-34
                            county_educ[,59]+ #Female: Grad/Prof Deg age 25-34
                            county_educ[,66]+ #Female: Bach Deg age 35-44
                            county_educ[,67]+ #Female: Grad/Prof Deg age 35-44
                            county_educ[,74]+ #Female: Bach Deg age 45-64
                            county_educ[,75]+ #Female: Grad/Prof Deg age 45-64
                            county_educ[,82]+ #Female: Bach Deg age 65+
                            county_educ[,83]), #Female: Grad/Prof Deg age 65+
                          denominator=county_educ[,1]) #Total Degrees

print (prop_bdegplus)
```


#Question #3: Download data for all of the zip codes in Fulton county on family income by zip code.
```{r Family Income by Zip code for Fulton County}
# 19001-- family income           
county_income<-acs.fetch(geography=select_zip, 
                        endyear = 2016,
                        table.number="B19001", 
                        col.names="pretty")

acs.colnames(county_income)
```


#Question #4: Compute the proportion of the population that has family income above 75,000.
```{r Proportion of Pop with Family Income above 75K}

prop_above_75<-divide.acs(numerator=(county_income[,13]+ #HH Income: $75K to 99,999
                            county_income[,14]+ #HH Income: $100K to $124,999
                            county_income[,15]+ #HH Income: $125K to $149,999
                            county_income[,16]+ #HH Income: $150K to $199,999
                            county_income[,17]), #HH INcome: $200K+
                          denominator=county_income[,1]) #HH Income Total
print (prop_above_75)

```

```{r Tibble and Naming}
# Convert to tibble
county_df<-tibble(substr(geography(county_educ)[[1]],7,11),
                       as.numeric(estimate(prop_bdegplus)),
                       as.numeric(estimate(prop_above_75)))


# Give it easy to use names
names(county_df)<-c("zip","college_educ","income_75")
save(county_df,file="dav.RData")

print(head(county_df))
```



#Question #5: Plot the proportion of residents with incomes above 75,000 as a function of income.
```{r Plot Proportion, echo=T, warning=F, message=F, fig.show='asis'}
gg<-ggplot(county_df,aes(x=college_educ,y=income_75))+
  geom_point(alpha=.5,size=.25)+
  scale_y_continuous(breaks = seq(10,50, by=5))+
  scale_color_discrete()+
  geom_smooth(method = "lm")+ # will default to the LOESS method
  geom_smooth(method = "loess",color="red")+
  geom_smooth(color="orange")+labs(title = "Proportion of Fulton County, GA Residents with Incomes Above $75K", x="College Education",y="Income")
print(gg)
```

