---
title: "10 - Assignment_BrundageK"
author: "Kelley Brundage"
affiliation: "Vanderbilt University - Peabody College, Ed.D. Leadership & Learning in Organizations program",
    "LLO 8200: Data Science"
date: "7/10/2019"
Github: "https://github.com/klbrundage/LLO8200-Assignments"
output: pdf_document
---

```{r setup, include=FALSE}
##This code allows the Knit function to still work even with errors 
knitr::opts_chunk$set(echo = FALSE, warning=FALSE, results ='hide',include=TRUE,messages=FALSE)

#We always start with a standard set of setup commands by loading the correct libraries. We will continue to work with our existing libraries and will add 'caret' in order to evaluate the perforance of a classifier.

##Load libraries in order to successfully run the code below - the suppressMessages coding will stop the install.packages information, etc.. from coming up in the Console and showing you what has run.

suppressMessages(library(caret)) #Misc functions for training and plotting classification and regression models.
suppressMessages(library(modelr)) #Functions for modelling that help you seamlessly integrate modelling into a pipeline of data manipulation and visualisation.
suppressMessages(library(tidyverse)) #set of packages that work in harmony because they share common data representations and 'API' design
suppressMessages(library(tictoc))#his package provides the timing functions 'tic' and 'toc' that can be nested. One can record all timings while a complex script is running, and examine the values later.

##Define My PDF setup
#This code does not show in the final document but will assist with definining the margin cutoff point and wraps the text to the next line.
knitr::opts_chunk$set(fig.path = "Figs/", results='hide', tidy.opts=list(width.cutoff=60)) 

  my_pdf = function(file,width,height)
  {pdf(file, width=width, height=height,pointsize=12)}

##Load the Dataset
#The code below will load the ELS Training and Testing Datasets used in Assignment #5


load("els_train.RData")
load("els_test.RData")
```

```{r Assignment 5 Code}
mod5 <- lm(bynels2r~byses1, data=els_train)
summary(mod5)
confint(mod5)

rmse_5 <- modelr::rmse(mod5, els_test);rmse_5

mod_6 <- lm(bynels2r~as.factor(byrace)+
              byses1,
            data=els_train)
summary(mod_6)

els_train <- els_train%>%
  add_predictions(mod_6)%>%
  rename(pred2=pred)

rmse_6 <- modelr::rmse(mod_6, els_test);rmse_6
```



Based on your work in assignment 5, do the following:

# *Question 1: Create a 10-fold cross validation of your linear model predicting reading scores as a function of at least two covariates. Provide a summary table or graphic of the RMSEs from this cross validation.


```{r Define & Run Model}
## Define the model

mod1_formula<-formula(read_rank~byses1+
                        byrace)
## Run the model against all of the data
basic.mod<-lm(mod1_formula,
              data=els_train) 

print(summary(basic.mod),
      only.contents=T,
      comment=F,
      sanitize.colnames.function=identity,
      sanitize.rownames.function=identity,
      hline.after=0:3)

```


The `crossv_kfold` command creates a list of datasets from our original dataset, each of which contains a testing and training dataset. The proportion of cases held out for testing is determined by the number of folds: 10 folds would indicate 1/10 of the data to be held out. 

```{r Kfold with 10 folds}

els_train_cf <- els_train%>%
  crossv_kfold(10)

print(els_train_cf,
      only.contents=T,
      comment=F,
      sanitize.colnames.function=identity,
      sanitize.rownames.function=identity,
      hline.after=0:3)
```


```{r Convert Training to Tibbles}

rmse_mod1 <- els_train_cf%>%
  mutate(train = map(train, as_tibble))%>% ##converts to tibbles
  mutate(model = map(train, ~lm(mod1_formula, data = .)))%>%
  mutate(rmse = map2_dbl(model, test, rmse))%>% ##apply model, get rmse
  select(.id, rmse) ##pull just id and rmse

rmse_mod1
```
Generate RMSE for Model1
```{r Model1 RMSE}
#The code below will print the RMSE calculated for Model 1

print(summary(rmse_mod1$rmse,
      only.contents=T,
      comment=F,
      sanitize.colnames.function=identity,
      sanitize.rownames.function=identity,
      hline.after=0:2))
```


The resulting dataset includes the id for the cross validation and the rmse. We can summarize and plot this new data frame to see what our likely range of rmse happens to be. 

```{r Plot RMSE dataframe, results="asis"}

gr1 <- ggplot(rmse_mod1, aes(rmse))+
  geom_density(bin=50, fill="sienna", alpha=.2)+
  labs(title = "Density Plot for RMSE Dataframe", x="RMSE", y="Density")

gr1
```


# *Question 2: Using a random partition, create 100 separate cross validations of your linear model predicting reading scores as a function of at least two covariates.  Provide a summary table or graphic of the RMSEs from this cross validation.

```{r Generalize Crossfold data}

els_train_cv <- els_train%>%
  crossv_mc(n=1000, test=.2)##prop of data to be held out - 20% or 1,000 rows to test

els_train_cv
```

The `els_train_cv` dataset is a dataset of 1000x2 datasets, with each row containing a training and testing dataset. The testing dataset is .2 or 20% of the sample, but it's different each time. 

```{r Model1 RMSE CV}
#Now we use the same approach, but with the MUCH larger qf_cv dataset. This will take a bit of time. 

mod1_rmse_cv<-els_train_cv %>% 
  mutate(train = map(train, as_tibble)) %>% ## Convert to tibbles
  mutate(model = map(train, ~ lm(mod1_formula, data = .)))%>%
  mutate(rmse = map2_dbl(model, test, rmse))%>% 
  select(.id, rmse) ## pull just id and rmse 

mod1_rmse_cv
```

```{r Summary of Model 1 RMSE CV}
print(summary(mod1_rmse_cv$rmse,
              only.contents=T,
              comment=F,
              sanitize.colnames.function=identity,
              sanitize.rownames.function=identity,
              hline.after=0:2))
```


```{r Summary & Plot of Model1 RMSE CV, results="asis"}



gr2 <- ggplot(mod1_rmse_cv, aes(rmse))+
  geom_density(bin=50, fill="thistle", alpha=.2)+
  labs(title = "Plot of Model 1 RMSE Cross Validation", x="RMSE", y="Density")

gr2
```
