---
title: "10-Assignment BrundageKelley"
author: "Kelley Brundage"
date: "7/17/2019"
Github: "https://github.com/klbrundage/LLO8200-Assignments"
output:
  pdf_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=FALSE,warning=FALSE,results='hide',include=TRUE,messages=FALSE)

##We always start with a standard set of setup commands by loading the correct libraries. We will continue to work with our existing libraries and will add 'caret' in order to evaluate the perforance of a classifier.

##Load libraries in order to successfully run the code below - the suppressMessages coding will stop the install.packages information, etc.. from coming up in the Console and showing you what has run.

suppressMessages(library(caret)) #Misc functions for training and plotting classification and regression models.
suppressMessages(library(dplyr)) #able to select, filter, organize, and manipulate data stored within an R data frame
suppressMessages(library(evaluate)) #Parsing and Evaluation Tools that Provide More Details than the Default
suppressMessages(library(forcats)) #Tools for Working with Categorical Variables (Factors)
suppressMessages(library(formatR)) #Provides a function tidy_source() to format R source code.
suppressMessages(library(ggplot2)) #A system for 'declaratively' creating graphics, based on "The Grammar of Graphics".
suppressMessages(library(haven)) #Import foreign statistical formats into R via the embedded 'ReadStat' C library
suppressMessages(library(knitr))#General-Purpose Package for Dynamic Report Generation in R 
  opts_chunk$set(comment = NA)
  def_hook <- knit_hooks$get("output")
  knit_hooks$set(output = function(x, options)
    {out <- def_hook(x, options)
    return(paste("\\begin{framed}\\begin{verbatim}", x, "\\end{verbatim}\\end{framed}",
                 collapse = "\n"))})

suppressMessages(library(lubridate)) #Functions to work with date-times and time-spans: fast and user friendly parsing of date-time data, extraction and updating of components of a date-time
suppressMessages(library(ModelMetrics)) #Collection of metrics for evaluating models written in C++ using 'Rcpp'.
suppressMessages(library(modelr)) #Functions for modelling that help you seamlessly integrate modelling into a pipeline of data manipulation and visualisation.
suppressMessages(library(pander))#provide a minimal and easy tool for rendering R objects
  panderOptions('table.style', "multiline")
  panderOptions('table.alignment.default',function(df)ifelse(sapply(as.data.frame(df),
                                                            is.numeric),'right','left'))

suppressMessages(library(readxl)) #reads in Excel Files
suppressMessages(library(rvest)) #scraping websites
suppressMessages(library(tibble)) #Provides a 'tbl_df' class (the 'tibble') that provides stricter checking and better formatting than the traditional data frame.
suppressMessages(library(tidyverse)) #set of packages that work in harmony because they share common data representations and 'API' design
suppressMessages(library(tictoc))#his package provides the timing functions 'tic' and 'toc' that can be nested. One can record all timings while a complex script is running, and examine the values later.

##Define My PDF setup
#This code does not show in the final document but will assist with definining the margin cutoff point and wraps the text to the next line.

knitr::opts_chunk$set(fig.path = "Figs/", results='hide', tidy.opts=list(width.cutoff=60)) 
  my_pdf = function(file,width,height)
  {pdf(file, width=width, height=height,pointsize=12)}

##Load the Dataset
#The code below will load the ELS Training and Testing Datasets used in Assignment #5

load("els_train.RData")
load("els_test.RData")
```

The summary results of Assignment #5 Code are below which is needed for the completion of Assignment #10 as it pulls from the ELS Train and ELSE Test datasets.  The code below is a compilation of a linear model (mod_5) of reading scores by Socioeconomic Status, coefficient of that lm, RMSE of that model and then a linear model  (mod_6) of reading scores as a factor of  race and Socioeconomic Status, predictions of the second model and the RMSE of the second model.  These are items needed in order to complete the homework assignment.

```{r Assignment 5 code}
##Pulled in code from Assignment 5 to help faciliate the Assignment 10 models

mod_5 <- lm(bynels2r~byses1, data = els_train)
summary(mod_5)
confint(mod_5)

rmse_5 <- modelr::rmse(mod_5, els_test);
rmse_5

mod_6 <- lm(bynels2r~as.factor(byrace)+
              byses1,
            data=els_train)
summary(mod_6)

els_train <- els_train%>%
  add_predictions(mod_6)%>%
  rename(pred2=pred)

rmse_6 <- modelr::rmse(mod_6, els_test);
rmse_6
```


## Question 1: 

### *Create a 10-fold cross validation of your linear model predicting reading scores as a function of at least two covariates. Provide a summary table or graphic of the RMSEs from this cross validation.*  


Define and Run the first Model which is from the ELS_Train dataset and is looking at reading scores compared to the two covariates of Socioeconomic Status and Race.
```{r Define & Run Model 1, results='asis'}
##The code below will define model 1 held within the els_train dataset 

mod1_formula <- formula(bynels2r~byses1+byrace)

basic.mod <- lm(mod1_formula,
                data = els_train)

pander(summary(basic.mod, justify="center"))
```

```{r Kfold with 10 folds}
##The `crossv_kfold` command creates a list of datasets from our original dataset, each of which contains a testing and training dataset. The proportion of cases held out for testing is determined by the number of folds: 10 folds would indicate 1/10 of the data to be held out.

els_train_cf <- els_train%>%
  crossv_kfold(10)

els_train_cf
```

ELS_Train converted to tibbles resulting in rmse for model 1:  
  
```{r Convert Training to Tibbles, results='asis'}
#the code below will convert the dataset els_train_cf to Tibbles and then apply the rmse to that model

rmse_mod1 <- els_train_cf%>%
  mutate(train = map(train, as_tibble))%>% ##converts to tibbles
  mutate(model = map(train, ~lm(mod1_formula, data = .)))%>%
  mutate(rmse = map2_dbl(model, test, rmse))%>% ##apply model, get rmse
  select(.id, rmse) ##pull just id and rmse

pander(rmse_mod1, caption="RMSE Apply Model 1")
```

  
The RMSE for Model 1 reflects:  

```{r Generate RMSE for Model 1, results="asis"}
##The code below will print the RMSE calculated for Model 1

pander(rmse_mod1$rmse, caption="RMSE for Model 1")

```


The resulting dataset includes the id for the cross validation and the RMSE We can summarize and plot this new data frame to see what our likely range of RMSE happens to be.  
```{r Plot RMSE datframe, fig.align="center",results='asis'}
##plot the rmse for model 1

gr1 <- ggplot(rmse_mod1, aes(rmse))+
  geom_density(bin=50, fill="sienna", alpha=.2)+
  labs(title = "Density Plot for RMSE Dataframe", x="RMSE", y="Density")

gr1
```


## Question 2:  
  
### *Using a random partition, create 100 separate cross validations of your linear model predicting reading scores as a function of at least two covariates.  Provide a summary table or graphic of the RMSEs from this cross validation.*    

```{r Generalize the Crossfold Data}
##the code below will generate a 100 separate cross fold validation models

els_train_cv <- els_train%>%
  crossv_mc(n=100, test=.2) ##proportion of data to be held out: 20% or 100 rows to test

els_train_cv
```

ELS_Train converted to tibbles resulting in rmse for model 1 cross validation:  
```{r RMSE Model 1 CV, results='asis'}
#Now we use the same approach, but with the MUCH larger els_train_cv dataset. 

mod1_rmse_cv <- els_train_cv%>%
  mutate(train = map(train, as_tibble))%>% ## Convert to tibbles
  mutate(model = map(train, ~lm(mod1_formula, data = .)))%>%
  mutate(rmse = map2_dbl(model, test, rmse))%>% ##apply model, get rmse cv
  select(.id, rmse) ## pull just id and rmse 

knitr::kable(summary(mod1_rmse_cv, caption="Cross Validation of RMSE Model 1"))
```

Within the cross validation of the RMSE Model 1 we see:
```{r Summary of RMSE Model 1 CV, results="asis"}

pander(summary(mod1_rmse_cv$rmse))
```

Below is a density plot for the RMSE Model 1 with cross validation:  
```{r Plot of RMSE Model 1 CV, fig.align="center",results="asis"}

gr2 <- ggplot(mod1_rmse_cv, aes(rmse))+
  geom_density(bin=50, fill="thistle", alpha=.2)+
  labs(title = "Plot of Model 1 RMSE Cross Validation", x="RMSE", y="Density")

gr2
```

Below is a density plot comparing the two cross validation models:  
```{r Plot to Compare CV Models, fig.align="center",results="asis"}

gr3 <- ggplot(mod1_rmse_cv, aes(x=rmse))+
  geom_density(fill="violet", alpha=.2)+
  geom_density(data=rmse_mod1,aes(x=rmse), fill="turquoise", alpha=.2)+
  labs(title = "Comparison of Cross Validation Models", x="RMSE", y="Density")

gr3
```
