---
title: "11-Assignment BrundageKelley"
author: "Kelley Brundage"
date: "7/24/2019"
Github: "https://github.com/klbrundage/LLO8200-Assignments"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=FALSE,warning=FALSE,results='hide',include=TRUE,messages=FALSE)

##We always start with a standard set of setup commands by loading the correct libraries. We will continue to work with our existing libraries and will add 'caret' in order to evaluate the perforance of a classifier.

##Load libraries in order to successfully run the code below - the suppressMessages coding will stop the install.packages information, etc.. from coming up in the Console and showing you what has run.

suppressMessages(library(caret)) #Misc functions for training and plotting classification and regression models.
suppressMessages(library(dplyr)) #able to select, filter, organize, and manipulate data stored within an R data frame
suppressMessages(library(evaluate)) #Parsing and Evaluation Tools that Provide More Details than the Default
suppressMessages(library(forcats)) #Tools for Working with Categorical Variables (Factors)
suppressMessages(library(formatR)) #Provides a function tidy_source() to format R source code.
suppressMessages(library(ggplot2)) #A system for 'declaratively' creating graphics, based on "The Grammar of Graphics".
suppressMessages(library(haven)) #Import foreign statistical formats into R via the embedded 'ReadStat' C library
suppressMessages(library(knitr))#General-Purpose Package for Dynamic Report Generation in R 
  opts_chunk$set(comment = NA)
  def_hook <- knit_hooks$get("output")
  knit_hooks$set(output = function(x, options)
    {out <- def_hook(x, options)
    return(paste("\\begin{framed}\\begin{verbatim}", x, "\\end{verbatim}\\end{framed}",
                 collapse = "\n"))})

suppressMessages(library(lubridate)) #Functions to work with date-times and time-spans: fast and user friendly parsing of date-time data, extraction and updating of components of a date-time
suppressMessages(library(ModelMetrics)) #Collection of metrics for evaluating models written in C++ using 'Rcpp'.
suppressMessages(library(modelr)) #Functions for modelling that help you seamlessly integrate modelling into a pipeline of data manipulation and visualisation.
suppressMessages(library(pander))#provide a minimal and easy tool for rendering R objects
  panderOptions('table.style', "multiline")
  panderOptions('table.alignment.default',function(df)ifelse(sapply(as.data.frame(df),
                                                            is.numeric),'right','left'))

suppressMessages(library(readxl)) #reads in Excel Files
suppressMessages(library(rvest)) #scraping websites
suppressMessages(library(tibble)) #Provides a 'tbl_df' class (the 'tibble') that provides stricter checking and better formatting than the traditional data frame.
suppressMessages(library(tidyverse)) #set of packages that work in harmony because they share common data representations and 'API' design
suppressMessages(library(tictoc))#his package provides the timing functions 'tic' and 'toc' that can be nested. One can record all timings while a complex script is running, and examine the values later.

##Special libraries used for Assignment #11
suppressMessages(library(nycflights13))
suppressMessages(library(RSQLite))

##Define My PDF setup
#This code does not show in the final document but will assist with definining the margin cutoff point and wraps the text to the next line.

knitr::opts_chunk$set(fig.path = "Figs/", results='hide', tidy.opts=list(width.cutoff=60)) 
  my_pdf = function(file,width,height)
  {pdf(file, width=width, height=height,pointsize=12)}

##Open the tables from Data sources from NYC flights
data(flights)
data(airlines)
data(airports)
data(weather)
data(planes)
```


# Question 1:  

## *Using the nyclfights13 library, open all of the data tables and turn them into a database. Make sure to include flights, airlines, airports, weather and planes.**  

```{r Connect to Databases}

con <- dbConnect(RSQLite::SQLite(), ":memory:")

##Write flights tables to database (you won't usually do this)
##con = connection - opening up a connection to the db

dbWriteTable(con,
             "flights", 
             as.data.frame(flights))

dbWriteTable(con,
             "planes", 
             as.data.frame(planes))

dbWriteTable(con,
             "airlines", 
             as.data.frame(airlines))

dbWriteTable(con,
             "weather", 
             as.data.frame(weather))

dbWriteTable(con,
             "airports", 
             as.data.frame(airports))
```

List of Tables in the Recently created Databases:  

```{r Table Access}
#List tables in database

pander(dbListTables(con))
```


List of Fields within each of the Databases:  

Airline Database Fields:  
```{r Airlines Database, results='asis'}
##List fields within airlines database

pander(dbListFields(con,"airlines"))
```

Airports Database Fields:  
```{r Airports Database, results='asis'}
##List fields within airports database

pander(dbListFields(con,"airports"))
```


Flights Database Fields:  
```{r Flights Database, results='asis'}
##List fields within flights database

pander(dbListFields(con,"flights"))
```

Planes Database Fields:  
```{r Planes Database, results='asis'}
##List fields within planes database

pander(dbListFields(con,"planes"))
```

Weather Database Fields:
```{r Weather Database Fields, results='asis'}
##List fields within weather database

pander(dbListFields(con,"weather"))
```


# Question 2:  

## *Create a data frame from a SQL pull from the database that consists only of flights that took off from JFK in May.*  

```{r Get text and turn into db}
#Generate a SQL request to a database
req_text <- "Select * from flights"

#Send query through connection
req <- dbSendQuery(con,req_text)

#Generate dataframe from results
req_df <- dbFetch(req,n=-1)

#Good practice: clear request
dbClearResult(req)

```

Flights Generated Dataframe:  
  
```{r Flights Generated Dataframe, results='asis'}
dim(req_df)

head(req_df,20)
```

Only flights that took off from JFK in the month of May (5) contained within the flights dataframe:  
  
```{r Flights Origin and Month, results='asis'}

req1_text <- "SELECT f.origin, f.month 
            FROM flights f 
            WHERE f.month=5 and f.origin='JFK'
            "
req1 <- dbSendQuery(con,req1_text)

req1_df <- dbFetch(req1,n=-1)

dbClearResult(req1)

head(req1_df)
```


# Question 3:  

## *Create a data frame from a SQL pull from the database that consists only of flights that took off on-time (a delay of less than 10 minutes) from Newark at temperatures of less than 40 degrees F.*  

```{r Flights from Newark, results='asis'}

req1_text <- "SELECT f.origin
              FROM flights f
              "

req1 <- dbSendQuery(con,req1_text)

req1_df <- dbFetch(req1,n=-1)

dbClearResult(req1)

head(req1_df) 
```


```{r Combine Flight with Weather, results='asis'}
#SQL request, flight info combined with weather info

weather_text <- "SELECT f.origin, f.dep_delay, w.temp
              FROM weather w
              JOIN flights f
              ON f.year=w.year
              AND f.month=w.month
              AND f.day=w.day
              WHERE f.origin='LGA' and f.dep_delay<10 AND w.temp<40"

weather_req <- dbSendQuery(con,weather_text)

weather_df <- dbFetch(weather_req,n=-1)

dbClearResult(weather_req)

head(weather_df)
```


# Question 4:  

## *Create data frame from a SQL pull from the database that consists of planes flown by United.*

```{r Airline data from Airline table}

req1_text <- "SELECT a.carrier, a.name
              FROM airlines a
              "

```

```{r Planes Flown by United, results='asis'}
req1_text<-"SELECT a.carrier, a.name, f.carrier, f.tailnum, p.tailnum 
            FROM airlines a, planes p
            JOIN flights f 
            ON f.tailnum=p.tailnum
            AND f.carrier=a.carrier
            WHERE a.carrier='UA'
            "
req1<-dbSendQuery(con,req1_text)

req1_df<-dbFetch(req1,n=-1)

dbClearResult(req1)

head(req1_df)
```
